Capstone Movielens Report

### 2025-11-26 reports/capstone-movielens.site/02.5.UME-model.Rmd
Report: UME Model documented.

::: {.noteblock data-latex=""}
In the code snippet above we use the [train_user_movie_effect.cv](#func.train_user_movie_effect.cv) function described in the [UME Model: Utility Functions] section of [Appendix A: Support Functions].
:::

@@

UM_effect.loop_starter <- c(endpoints["start"], 
                            endpoints["end"], 
                            8)
UME.rglr.fine_tune.cache.base_name <- "UME.rglr.fine-tune"
```
```{r}
print(UM_effect.loop_starter)
print(UME.rglr.fine_tune.cache.path)
print(UME.rglr.fine_tune.cache.base_name)
class(regularize.test_lambda.UM_effect.cv)
```
```{r, eval=FALSE}

@@

UME.rglr.fine_tune.RMSE.best <- UME.rglr.fine_tune.results$best_result["best_RMSE"]
```
```{r, echo=FALSE}
writeLines("*** Fine-tuning results object data structure ***")
str(UME.rglr.fine_tune.results)

writeLines("*** Fine-tuning: best results ***")
UME.rglr.fine_tune.results$best_result
```

Let's visualize the fine-tuning results:
```{r}
UME.rglr.fine_tune.results$tuned.result |>
  data.plot(title = "UME Model Regularization: Fine-tuned result",
              xname = "parameter.value",
              yname = "RMSE",
              xlabel = TeX(r'[$\lambda$]'),
              ylabel = str_glue("Deviation from the best RMSE value (",
                                as.character(round(UME.rglr.fine_tune.RMSE.best, digits = 7)), 
                                ")"),
              normalize = TRUE)
```
\newpage

@@

Now, we can calculate the _Regularized User+Movie Effect_ by retraining our model on the entire `edx` dataset with the best value of the $\lambda$ parameter we just calculated, for the definitive _Root Mean Squared Error_ calculation and use in subsequent models.
```{r, eval=FALSE}
  UME.rglr.best_lambda <- best_result["param.best_value"]
  UME.rglr.best_lambda

  rglr.UM_effect <- train_user_movie_effect(edx, UME.rglr.best_lambda)
```
```{r, echo=FALSE}
writeLines("*** The Best Fine-tuning Results ***")
UME.rglr.fine_tune.results$best_result

writeLines("*** Regularized User+Movie Effect Structure ***")
str(rglr.UM_effect)

@@

::: {.noteblock data-latex=""}
The complete version of the source code provided in this section are available in the [Re-training Regularized User+Movie Effect Model for the best $\lambda$](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L910) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L910) script on _GitHub_.
:::

We calculate the _Root Mean Squared Error_ for the ultimately computed _User+Movie Effect_ using [calc_user_movie_effect_RMSE.cv](#func.calc_user_movie_effect_RMSE.cv) function described above as follows:
```{r, eval=FALSE}
UME.rglr.retrain.RMSE <- calc_user_movie_effect_RMSE.cv(rglr.UM_effect)
```
```{r}
print_log1("The best RMSE after being regularized: %1",
         UME.rglr.retrain.RMSE)
```

@@

::: {.noteblock data-latex=""}
Here use the [RMSEs.AddRow](#func.RMSEs.AddRow) and [RMSE_kable](#func.RMSE_kable) functions described in the [Result RMSEs Tibble Functions] section of [Appendix A: Support Functions].
:::

>>>>

::: {.noteblock data-latex=""}
In the code snippet above we use the [train_user_movie_effect.cv](#func.train_user_movie_effect.cv) function described in the *Models Training: Support Functions /* [UME Model: Utility Functions] section of [Appendix A: Support Functions].
:::

@@

UM_effect.loop_starter <- c(endpoints["start"], 
                            endpoints["end"], 
                            8)
UME.rglr.fine_tune.cache.base_name <- "UME.rglr.fine-tune"
```

```{r}
print(UM_effect.loop_starter)
print(UME.rglr.fine_tune.cache.path)
print(UME.rglr.fine_tune.cache.base_name)
```
```{r, eval=FALSE}

@@

UME.rglr.fine_tune.RMSE.best <- UME.rglr.fine_tune.results$best_result["best_RMSE"]
```
```{r, echo=FALSE}
writeLines("*** Fine-tuning results object data structure ***")
```
```{r}
str(UME.rglr.fine_tune.results)
```
```{r, echo=FALSE}
writeLines("*** Fine-tuning: best results ***")
```
```{r}
UME.rglr.fine_tune.results$best_result
```

Let's visualize the fine-tuning results:
```{r}
UME.rglr.fine_tune.results$tuned.result |>
  data.plot(title = "UME Model Regularization: Fine-tuned result",
              xname = "parameter.value",
              yname = "RMSE",
              xlabel = TeX(r'[$\lambda$]'),
              ylabel = str_glue("Deviation from the best RMSE value (",
                                as.character(round(UME.rglr.fine_tune.RMSE.best, digits = 7)), 
                                ")"),
              normalize = TRUE)
```

::: {.noteblock data-latex=""}
The complete version of the source code provided in this section are available in the [Fine-tuning Step of the Regularization Method for the User+Movie Model](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L607) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L607) script on _GitHub_.
:::

@@

Now, we can refine our _User+Movie Effect Model_ by retraining on the entire `edx` dataset with the best value of the $\lambda$ parameter we just figured out (let's call it *Regularized User+Movie Effect Model*), for the definitive _Root Mean Squared Error_ calculation and use in subsequent models.
```{r, eval=FALSE}
  UME.rglr.best_lambda <- best_result["param.best_value"]
  UME.rglr.best_lambda

  rglr.UM_effect <- train_user_movie_effect(edx, UME.rglr.best_lambda)
```

::: {.noteblock data-latex=""}
In the code snippet above we use the [train_user_movie_effect](#func.train_user_movie_effect) function described in the *Models Training: Support Functions /* [UME Model: Utility Functions] section of [Appendix A: Support Functions].

Later in this section, we will also use the custom logging function [print_log1](#func.print_log1) described in the [Logging Functions] section of *the same Appendix*.
:::

```{r, echo=FALSE}
writeLines("*** The Best Fine-tuning Results ***")
```
```{r}
UME.rglr.fine_tune.results$best_result
```

```{r, echo=FALSE}
writeLines("*** Regularized User+Movie Effect Structure ***")
```
```{r}
str(rglr.UM_effect)

@@

::: {.noteblock data-latex=""}
The complete version of the source code provided in this section are available in the [Re-training Regularized Model for the best `lambda`](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L650) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L650) script on _GitHub_.
:::

We calculate the *RMSE* for the ultimately *Regularized User+Movie Effect Model* using [calc_user_movie_effect_RMSE.cv](#func.calc_user_movie_effect_RMSE.cv) function already used above in the [UME Model Building] section:
```{r, eval=FALSE}
UME.rglr.retrain.RMSE <- calc_user_movie_effect_RMSE.cv(rglr.UM_effect)
```
```{r}
print_log1("The best RMSE after being regularized: %1",
         UME.rglr.retrain.RMSE)
```

@@

::: {.noteblock data-latex=""}
To print the table above, we use the [RMSEs.AddRow](#func.RMSEs.AddRow) and [RMSE_kable](#func.RMSE_kable) functions described in the [Result RMSEs Tibble Functions] section of [Appendix A: Support Functions].
:::
